Prediction Markets for Exponential Families

### 2014-01-31 WriteUp/Original/maxlike-scoring.tex
tidied up main folder to just have submission doc

% Adjusted to add \usepackage[numbers]{natbib}
% \bibliographystyle{acmsmall} and \documentclass[prodmode,acmec]{ec-acmsmall}
% Jan 5, 2013 - David Parkes
%
% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmec]{ec-acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2013}
\acmMonth{6}
\usepackage[numbers]{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{parskip}
\usepackage{booktabs}

%new commands
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\bR}{\mathbf{R}}
\newcommand{\grad}{\nabla}
\newcommand{\Exp}{\mathbf{E}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\q}{\mathbf{q}}
\newcommand{\mM}{\ensuremath{\mathcal{M}}}
\newcommand{\hmu}{\ensuremath{\hat{\mu}}}
\newcommand{\mP}{\ensuremath{\mathcal{P}}}
\newcommand{\mY}{\ensuremath{\mathcal{Y}}}
\newcommand{\mS}{\ensuremath{\mathcal{S}}}
\newcommand{\ds}{\displaystyle}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\htheta}{\hat{\theta}}
%---------------------------------------------------------------------------
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{con}[thm]{Conjecture}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{ex}[thm]{Example}
\newtheorem{qn}[thm]{Question}
\newcommand{\defn}{\bigbreak\noindent{\bf Definition. }}
\newcommand{\undefn}{\bigbreak}
\newcommand{\rmk}{\bigbreak\noindent{\bf Remark. }}
\newcommand{\unrmk}{\bigbreak}
%\newcommand{\proof}{\noindent{\bf Proof. }}
\newcommand{\unproof}{\hfill$\Box$\bigbreak}
\newcommand{\defeq}{\stackrel{\mathrm{def}}{=}}
\newcommand{\holds}[1]{\stackrel{?}{#1}}
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\ie} {{\it i.e. }}
\newcommand{\eg} {{\it e.g. }}
\newcommand{\etal} {{\it et al. }}
\newcommand{\E}{\mathbf{E}}
\newcommand{\betavec}{\pmb{\beta}}
\newcommand{\qvec}{\mathbf{q}}
\newcommand{\lpf}{\mathbf{\psi}} %logpartition function
\newcommand{\family}{\mathcal F} 
\newcommand{\suff}{\pmb{\phi}} % suff stats
\newcommand{\muvec}{\pmb{\mu}}

% Document starts

\begin{document}
% Page heads
\markboth{Abernethy et al.}{Maximum Entropy Prediction Markets}
% Title portion
\title{Maximum Entropy Prediction Markets}
\author{JACOB ABERNETHY
\affil{University of Michigan, Ann Arbor}
SINDHU KUTTY
\affil{University of Michigan, Ann Arbor}
S\'{E}BASTIEN LAHAIE
\affil{Microsoft Research, New York City}
RAHUL SAMI
\affil{University of Michigan, Ann Arbor}}
%%\author{GANG ZHOU
%%\affil{College of William and Mary}
%%YAFENG WU
%%\affil{University of Virginia}
%%TING YAN
%%\affil{Eaton Innovation Center}
%%TIAN HE
%%\affil{University of Minnesota}
%%CHENGDU HUANG
%%\affil{Google}
%%JOHN A. STANKOVIC
%%\affil{University of Virginia}
%%TAREK F. ABDELZAHER
%%\affil{University of Illinois at Urbana-Champaign}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets. We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data. Similarly,  the market state can be interpreted as a distribution over the outcome space. In particular, we show that a proper scoring rule can be derived from maximum entropy distributions. This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces. In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility. We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function. These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders. In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market. Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
\end{abstract}

\category{J.4}{Social and Behavioral Sciences}{Economics}
\category{I.2.6} {Artificial Intelligence}{Learning}

\terms{Prediction Market Design, Machine Learning Algorithms}

\keywords{Exponential Families, Bayesian Learning, Maximum Entropy Distributions, generalized LMSR}

\acmformat{Jacob Abernethy, Sindhu Kutty, S\'{e}bastien Lahaie, and Rahul Sami, 2014. Maximum Entropy Prediction Markets}
%%Gang Zhou, Yafeng Wu, Ting Yan, Tian He, Chengdu Huang, John A. Stankovic, and Tarek F. Abdelzaher, 2010. A multifrequency MAC specially designed for  wireless sensor network applications.

% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
%%This work is supported by the National Science Foundation, under
%%grant CNS-0435060, grant CCR-0325197 and grant EN-CS-0329609.
%%
%%Author's addresses: G. Zhou, Computer Science Department,
%%College of William and Mary; Y. Wu  {and} J. A. Stankovic,
%%Computer Science Department, University of Virginia; T. Yan,
%%Eaton Innovation Center; T. He, Computer Science Department,
%%University of Minnesota; C. Huang, Google; T. F. Abdelzaher,
%%(Current address) NASA Ames Research Center, Moffett Field, California 94035.
\end{bottomstuff}

\maketitle

\section{Introduction}
%------------------------------------------------------------
%Outline
%------------------------------------------------------------
%- exp family review
%- maxent scoring rule
%	- generalized lmsr
%	- does modularity (a la Hanson) still hold?
%- prediction market setup
%	- MLE with risk neutrality
%- conjugate prior market
%	- also on hyperparam?
%- exp-util
%- equilibrium for repeated trades
%- budgets and adversarial markets

We have dual goals in this paper. On the one hand, we highlight a structural similarity between prediction markets and exponential families. We see this as the syntax of our prediction market mechanism. On the other, this formulation has rich semantics as well: it allows for analysis of market behavior under various environments. For instance, we analyze market behavior with budget limited traders, traders that are exponential-utility-maximizers.
\section{Notation and Definitions}
convex conjugates, bregman divergences, scoring rules, sufficient statitics, exp families, Bayesian updates, conjugate prior.
\section{Maximum Likelihood Scoring Rules}

@@

\begin{itemize}
\item Explain that the scoring rule from Example 2 does not elicit the mean when the agent's distribution is exponential (with change of variable so that it lies in $[1,+\infty)$. 
\item Observe that scoring rule from Example does elicit the mean when the agent's distribution is exponential, Pareto, or in fact any other distribution with support on $[0,+\infty)$ such as the lognormal. 
\item This occurs because the scoring rule is linear in $x$, so its expectation is the same for any distribution with mean $\mu$. 
\item Another feature to note is that under an appropriate parametrization (rate instead of mean), the scoring rule (1) is convex in the parameter being elicited. 
\item These two observations taken together lay the groundwork for the characterization of strongly proper scoring rules for distribution statistics. 
\item Before we proceed should note that the arguments behind Theorem~\ref{maxlike} apply not just to maximum likelihood but to any \emph{m-estimator}, which are notions from robust statistics. Explain what an m-estimator looks like and how there are general consistency results for those as well.
\item The point of these alternative estimators is to be more robust to outliers in the sample, and so these kinds of estimators could translate into scoring rule more robust to inaccuracies in the agent's assessment of the underlying probability distribution or its parameters. (Just a hunch.)
\end{itemize}
\section{Maximum Entropy Market Making}

The purpose of a prediction market is to elicit and aggregate the beliefs (i.e., subjective probabilities) of agents over a space of \emph{outcomes}. In a single-agent setting, a scoring rule is used to elicit the agent's beliefs. In a multi-agent setting, an information market is used to aggregate the agent's beliefs. Hanson~\cite{} introduced the idea of a market scoring rule, which inherits the appealing elicitation and aggregation properties of both so that they can perform well in both thin and thick markets. In this work we derive a wide-ranging generalization of the logarithmic market scoring rule. Using a maximum entropy approach, we explain how a market scoring rule can be developed for outcome spaces both discrete and continuous, and for generic properties of the underlying distribution (e.g., mean and variance), rather than just the probabilities of individual outcomes.

\subsection{The Model}
\label{model}

Let $\mX$ be the outcome space, which may be discrete or continuous. Let $\mP$ denote the set of probability distributions over the outcome space. We represent a probability distribution as a density $p$ absolutely continuous with respect to some base measure $\nu$ (e.g., counting for discrete outcomes, or Lebesgue for continuous outcomes).% We will discuss the interpretation of the base measure in a later section.

There is an unknown distribution $p$ over outcomes. We are interested in estimating the expected value of different outcome \emph{statistics} under this distribution by aggregating the beliefs of agents. A statistic is a real-valued function $\phi_s : \mX \rightarrow \bR$ over outcomes, independent of $p$; here $s$ belongs to a finite index set $\mS$ of size $d$. The collection of functions $\phi = (\phi_s,\, s \in \mS)$ can be viewed as a vector-valued statistic mapping outcomes into $\bR^d$. Formally, the aim is to estimate
%
\begin{equation} \label{constraint}
\Exp_p[\phi(x)] = \mu.
\end{equation}
%
We stress that we are only seek to elicit $\mu \in \bR^d$, not the full distribution $p$. As an example, suppose that $\mX = \bR$. If we were interested in eliciting the first two uncentered moments of $p$, we would include $\phi_1(x) = x$ and $\phi_2(x) = x^2$ as statistics. Note that the variance cannot be directly obtained through any single statistic: it corresponds to the expectation of $(x - \Exp_p[x])$, which depends on $p$, violating our definition of a statistic. Of course, the variance can be indirectly obtained by eliciting the first two moments separately.\footnote{This is related to the notion of \emph{elicitation complexity} of statistical properties introduced by Lambert et al.~\cite{Lambert}, but we do not pursue this connection here.}

There are two approaches to eliciting an agent's estimated $\mu$ in~(\ref{constraint}). The first approach is to incentivize the agent to directly report $\mu$ via a \emph{scoring rule}. The second approach is to form an \emph{information market} by issuing securities for each statistic $s \in \mS$, with payoff depending on the realized value of the statistic. (Such securities are known as contingent claims.) The amount of shares of each security acquired by the agent indirectly reveals it estimate $\mu$. In a market scoring rule~\cite{Hanson}, these two approaches are dual to each other in a formal sense.

\subsection{Scoring Rule}

We first consider how to directly elicit $\mu$ via a scoring rule. 
%Note that the usual definition of a scoring rule presumes that a full probability distribution is elicited, rather than just some of its properties (e.g., mean and variance), so below we will adapt the definition to our setting. 
The space of feasible reports from an agent corresponds to
%
\begin{equation} \label{marginal}
\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}.
\end{equation}
%
A scoring rule is a function $S : \mM \times \mX \rightarrow \bR$ that rewards an agent with $S(\mu,x)$ based on how its report $\mu$ agreed with the eventual outcome $x$. A scoring rule is \emph{proper} if for all $\mu \in \mM$, and $p \in \mP$ such that $\Exp_p[\phi(x)] = \mu$, we have
%
\begin{equation} \label{scoring}
\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
\end{equation}
%
for all $\hmu \neq \mu$; the rule is \emph{strictly proper} if the inequality is strict. Note that in the literature, scoring rules are defined over entire probability distributions, not just properties of those distributions. Our definition imposes a minimum of informational requirements on agents, because an agent does not need to assess the full distribution $p$ that might have lead to its estimates $\mu$ in order to realize that truthful reporting is an optimal strategy.

Our scoring rule is implicitly defined via the solution of a mathematical program. The program computes the maximum entropy distribution consistent with the agent's reported beliefs $\mu$.
%
\begin{eqnarray}
\max_{p \geq 0} & \ds - \int_{x \in \mX} p(x) \log p(x) \nu(dx) & \label{obj} \\
\mbox{s.t.} & \ds \int_{x \in \mX} \phi(x) p(x) \nu(dx) = \mu & \label{mean} \\
 & \ds \int_{x \in \mX} p(x) \nu(dx) = 1 & \label{normalize}
\end{eqnarray}
%
The non-negativity constraints together with~(\ref{normalize}) ensure that the solution is a probability distribution. Constraints~(\ref{mean}), which correspond to $d$ separate constraints, ensure consistency with the agent's report. The objective~(\ref{obj}) is the entropy of the solution $p$ with respect to the measure $\nu$. The program has a convex objective and linear constraints. %For continuous outcome spaces we will see in the next section that the optimal solution can still be cleanly characterized. 
%
\begin{theorem} \label{thm:scoring}
Let $p(x;\mu)$ be the optimal solution to the maximum entropy program, given report $\mu \in \bR^d$. The scoring rule defined by
%
\begin{equation} \label{scoring}
S(\mu,x) = a_x + b\log p(x;\mu),
\end{equation}
%
where $b,a_x \in \bR$ and $b > 0$, is strictly proper.
\end{theorem}
%
Recall that a  proper scoring rule $\mathbf{S}$ satisfies $$ \E_{x\sim p}[\mathbf{S}(p,x)]\geq \E_{x\sim p}[\mathbf{S}(p^{\prime},x)]$$

Note that the maximum entropy distribution is an exponential family distribution. 
 Let $p(x)=\exp\{\beta\cdot \phi(x)-\psi(\beta)\}$ be the true distribution and  $p'(x)=\exp\{\beta'\cdot \phi(x)-\psi(\beta')\}$ be the predicted distribution on the event $x$.

 We need to show that $$\E_{x\sim p}[\mathbf{S}(p,x)]\holds{\geq} \E_{x\sim p}[\mathbf{S}(p',x)]$$
  In other words, we need $$\E_{x\sim p}[\log p(x)]\holds{\geq} \E_{x\sim p}[\log p'(x)]$$
   Equivalently, $$\E_{x\sim p}[\beta\cdot \phi(x)-\psi(\beta)]\holds{\geq} \E_{x\sim p}[\beta'\cdot \phi(x)-\psi(\beta')]$$
   Or 
\begin{equation}\label{eq:proper}
\beta\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta)\holds{\geq} \beta'\cdot \E_{x\sim p}[\phi(x)]-\psi(\beta')
\end{equation}

   For an exponential family distribution, $\E_{x\sim p}[\phi(x)]=\nabla\psi(\beta)$. Substituting in  $(\ref{eq:proper})$, we need to show
   $$\beta\cdot \nabla\psi(\beta)-\psi(\beta)\holds{\geq} \beta'\cdot \nabla\psi(\beta)-\psi(\beta')$$
   Rearranging the terms,
   $$(\beta-\beta')\cdot \nabla\psi(\beta)\holds{\geq} \psi(\beta)-\psi(\beta')$$   
   which is the first order condition on the convexity of $\psi(\cdot)$.
 \unproof

 \rmk The distance between the scores of the true and any predictive exponential family distribution is, in fact,  the Bregman divergence (based on the log partition function) between the corresponding natural parameters. That is,
   $$\E_{x\sim p}[\mathbf{S}(p,x)]- \E_{x\sim p}[\mathbf{S}(p',x)] = D_{\psi}(\beta,\beta')$$
   \unrmk

\noindent
This rule represents a generalization of the logarithmic scoring rule for probability distributions to any properties of distributions that can be captured as expectations of statistics.  As explained in the next section, the solution to the maximum entropy program takes the form of an \emph{exponential family} distribution. For many common properties of interest, these distributions are familiar; for example, the maximum entropy distribution with a given mean is an exponential distribution, and the maximum entropy distribution with a given mean and variance is a normal distribution~\cite{}. We provide here two examples in more depth.

%%
%%

\paragraph{Multinomial Distribution}

As a first example, consider the problem of estimating the individual probabilities of a finite set of outcomes. We have $\mX = \{1,\ldots,k\}$. The relevant statistics indicate which outcome actually occurs, so the index set is $\mS = \mX$. Statistic $\phi_x(x')$ is 1 if $x' = x$ and 0 otherwise. Observe that if $p$ is the distribution over outcomes, then $\Exp_p[\phi(x)] = p$. A feasible report from an agent is any non-negative $\pi = (\pi_1,\ldots,\pi_k)$ such that $\sum_{i=1}^k \pi_i = 1$. Given a report of $\mu = \pi$, the unique solution to the maximum entropy program is $p(x; \pi) = \pi$, and~(\ref{scoring}) corresponds to the classic logarithmic scoring rule of $S(\pi,x) = \log \pi_x$. \hfill{$\Box$}

\paragraph{Normal Distribution}

As a second example, suppose the outcome space is $\mX = \bR$ and that we are interested in estimating the mean $\mu$ and variance $\sigma^2$ of the underlying distribution. We choose the first and second (uncentered) moments, $\kappa_1 = \mu$ and  $\kappa_2 = \mu^2 + \sigma^2$ as our statistics.  It is well-known that the maximum entropy distribution with a given mean and variance is the normal distribution, so the solution to the maximum entropy program (using Lebesgue as the base measure $\nu$) is
%
\[ p(x; \kappa) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp\left[ -\frac{(x-\mu)^2}{2\sigma^2} \right]. \]
%
Applying Theorem~\ref{thm:scoring} and choosing appropriate constants, we find that the following scoring rule is strictly proper in this context. %
\[ S(\kappa, x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2. \]
%
Note that the fact that this scoring rule is proper for the mean amounts to the well-known fact that reporting the mean is a Bayes act under quadratic loss~\cite{someone}. \hfill{$\Box$}

\medskip\noindent
A brief observation on parametrizations is in order. Note that the maximum entropy distribution here is parametrized by the uncentered moments $\kappa = (\kappa_1,\kappa_2)$ rather than $(\mu,\sigma^2)$ because the variance is not a valid statistic. However, it is clear that the agent does not have to directly report $\kappa$: the market maker itself can do the translation from reported mean and variance to uncentered moments. This is relevant because the $\kappa$ parametrization may be unintuitive for an agent, especially given the constraints $\kappa_2 - \kappa_1^2 \geq 0$ that must be enforced. In the $(\mu, \sigma^2)$ parametrization each parameter is unrestricted, and the parameters have intuitive interpretations.

Our approach so far has been to identify relevant properties of the unknown distribution, and elicit them via a scoring rule computed through a maximum entropy program. However, it is the case that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}. Therefore, we see the following alternative approach. The market designer can begin by considering the outcome space $\mX$, and assessing what family of distributions over $\mX$ the unknown $p$ might come from. If this family is an exponential family, then Theorem~\ref{thm:scoring} immediately gives a proper scoring rule to elicit the parameters of the distribution. Many of the most familiar distributions are exponential families, including the normal, exponential, gamma, beta, binomial, Poisson, Weibull, and Dirichlet. To illustrate, we consider two examples in more depth.

%%
%%

\paragraph{Exponential}
Suppose the designer would like an estimate of the rate at which servers fail in a data center. An outcome is the time between failures, so $\mX = [0,+\infty)$. The usual distribution for failure times is the exponential distribution $p(x;\lambda) = \lambda e^{-\lambda x}$, which is parametrized by the rate $\lambda$ and has mean $\mu = 1/\lambda$. The exponential distribution is the maximum entropy distribution given a fixed mean $\mu$. From Theorem~\ref{thm:scoring} we obtain the following proper scoring rule for this setting:
%
\[ S(\mu, x) = \log \lambda - \lambda x. \]
%
It is simple to check using basic calculus that an agent maximizes the expected score by reporting $\lambda = 1/\hmu$, where here $\hmu$ is its estimate of the mean. In this case, it is unclear a priori which of the mean or rate parametrizations is most intuitive, but the maximum entropy approach easily allows for either. \hfill{$\Box$}

%%
%%

\paragraph{Beta}
Suppose the designer would like an estimate of the click-through rate of an online advertisement. The outcome space here is $\mX = (0,1)$. A typical prior over a probability like this is the beta distribution, which has two parameters $\alpha, \beta > 0$ and density function $p(x; \alpha, \beta) = x^{\alpha-1}(1-x)^{\beta-1}/B(\alpha,\beta)$, where $B$ is the beta function. By theorem~\ref{thm:scoring} this leads to the proper scoring rule
%
\[ S((\alpha, \beta), x) = (\alpha-1)\log x + (\beta-1)\log (1-x) - \log B(\alpha,\beta). \]
%
Although this parametrization is standard for the beta, it is perhaps more intuitive to use the parametrization $\mu = \alpha/(\alpha+\beta) \in (0,1)$ and $n = \alpha+\beta > 0$. Here $\mu$ is the mean of the distribution---corresponding to the agent's actual estimate of the click-through rate---and $n$ is a measure of the agent's confidence in the estimate. \hfill{$\Box$}

\subsection{Information Market}

We next consider how to indirectly elicit $\mu$ by setting up a market of contingent claim securities. Under this approach, the elements of the index set $\mS$ are interpreted not as statistics but as securities. The payoff from one share of security $s$ when outcome $x$ occurs is given by the mapping $\phi_s$. Thus if the vector of shares held by the agent is $\theta \in \bR^d$, where entry $\theta_s$ corresponds the number of shares of security $s$, then the payoff to the agent when $x$ occurs is evaluated by taking the inner product $\la \theta, \phi(x) \ra$. As a concrete example, recall our multinomial distribution example from the previous section, where $\phi_x(x') = 1$ if $x' = x$ and 0 otherwise. This means that security $x \in \mS$ pays 1 dollar if outcome $x \in \mX$ occurs, and nothing otherwise. (Such securities are known as Arrow-Debreu securities.) In our normal distribution example, the statistic for the mean had the mapping $\phi_1(x) = x$. Therefore a share of the corresponding security has a payoff that is linear in the outcome. (Such securities amount to futures contracts.)

To extract useful information from such a securities market, we set a pricing scheme and examine the number of shares the agent chooses to acquire. Let $\Omega$ be the set of all portfolios (i.e., vectors of shares) that the agent can feasibly hold; we will see how this domain is determined in an instant. Each security $s$ has an associated price function $c_s : \Omega \rightarrow \bR$, which gives the marginal price $c_s(\theta)$ of security $s$ when the agent holds portfolio $\theta$. (Note that the marginal price depends on the entire portfolio $\theta$, not just the number of shares $\theta_s$.) A risk-neutral agent will choose to acquire shares up to the point where, for each share, expected payoff equals marginal price. Formally, if the agent acquires portfolio $\theta$, then for each $s \in \mS$ we must have
%
\begin{equation} \label{market}
\Exp_p[\phi_s(x)] = c_s(\theta).
\end{equation}
%
In this way, by its choice of $\theta$, the agent reveals that its belief is $\mu = c(\theta)$. There are several important properties that the price function $c$ should have. To simplify the agent's portfolio acquisition process, it should not be the case that the total cost of acquiring $\theta$ depends on the order in which shares are bought. This means that there should exist a \emph{cost function} $C : \Omega \rightarrow \bR$ such that $c = \nabla C$. The gradient $\nabla C$ must be onto $\mM$ to ensure that~(\ref{market}) always has a solution---this is the most important but technical condition to realize. To ensure that the solution is unique, it would also be convenient if the cost function were strictly convex. 
%
\begin{proposition}
The following cost function is monotone, convex, and onto $\mM$: 
%
\begin{equation} \label{cost}
C(\theta) = \log \int_{x \in \mX} \exp \left[\la \theta, \phi(x) \ra\right] \nu(dx).
\end{equation}
%
\end{proposition}
%
\noindent
Observe that in the context of our earlier multinomial example, cost function~(\ref{cost}) is exactly the cost function for Hanson's logarithmic market scoring rule. Our approach here provides a generalization of this market scoring rule to markets with contingent claim securities with arbitrary payoffs, designed to elicit specific properties of distributions, beyond just the probabilities of different outcomes.

Because an agent would never select a portfolio with infinite cost, the effective domain of $C$ is $\Omega = \{\theta \:|\: C(\theta) < +\infty \}$. In the context of the multinomial distribution example, $\Omega = \bR^d$, so that shares of each security can always be bought or sold short. In the context of the normal distribution example the effective domain is $\Omega = \{(\theta_1,\theta_2) \in \bR^2 \:|\: \theta_2 > 0 \}$ if we re-define the second statistic to be $\phi_2(x) = -x^2$; this means that the security has a negative payoff for each share, and consequently the agent must be compensated to acquire such shares. Evaluating~(\ref{cost}), the cost function takes the form $C(\theta) = \frac{\theta_1^2}{4\theta_2} - \frac{1}{2}\log(2\theta_2)$.

Now, in our previous elicitation approach that used a scoring rule, the process of computing the scoring rule also provided a complete distribution $p(x;\mu)$ over outcomes, which could be used to infer other properties beyond just the agent beliefs $\mu$. In the current market-based approach, the agent's chosen portfolio $\theta$ can also form the basis of a distribution over outcomes. Consider the following distribution, represented as a density with respect to a base measure $\nu$:
%
\begin{equation} \label{expfam}
p(x;\theta) = \exp\left[ \la \theta, \phi(x) \ra - C(\theta) \right].
\end{equation}
%
Meaning, the probability that a subset $X \subseteq \mX$ of the outcomes occurs is $\int_{x \in X} p(x;\theta)\nu(dx)$. Observe that by definition~(\ref{cost}), this probability density indeed integrates to 1 over $\mX$, and it is clear that the density is non-negative as required.

In the statistical literature a distribution that takes the form~(\ref{expfam}) is known as an \emph{exponential family}. The mapping $\phi$ is known as the \emph{sufficient statistic}, $\theta$ is the \emph{natural parameter}, and $C$ is the log-partition or \emph{cumulant} function. %This connection between information markets and exponential families allows us to draw on a vast statistical literature to understand the market's properties, and a growing computer science literature to apply efficient algorithms to compute costs and perform inference.

\subsection{Duality}

There is a well-known duality between the maximum entropy approach and exponential families, which translates into a duality between the scoring rule and information market just developed. The duality implies that the approach leads to a \emph{market scoring rule}, applicable to both thin and thick multi-agent settings.

It is known that a distribution is a maximum entropy distribution if and only if it is an exponential family~\cite{}. To see this, let $\theta(\mu) \in \bR^d$ be the Lagrange multiplier corresponding to constraints~(\ref{mean}) when solving the maximum entropy program given the agent report $\mu$. (Our choice of notation is deliberately suggestive.) Let $A(\mu)$ be the Lagrange multiplier corresponding to~(\ref{normalize}). From the first order necessary conditions for optimality, we find that
%
\[ p(x;\mu) = \exp \left[ \la \theta(\mu), \phi(x) \ra - A(\mu) - 1 \right], \]
%
so that the solution is indeed in exponential family form. Conversely, given a cumulant $C$ from an exponential family and a parameter $\theta$, we see that $p(x;\theta)$ defined according to~(\ref{expfam}) is the maximum entropy distribution under constraints $\Exp_p[\phi(x)] = \mu$ where $\mu = \nabla C(\theta)$. We obtain the following.
%
\begin{proposition}
Let $C$ be the cumulant (i.e, cost function) for the exponential family corresponding to $\phi$, and let $\theta(\mu)$ be the optimal Lagrange multiplier for the mean constraints given an agent report of $\mu \in \mM$. Then the following scoring rule is equivalent to~(\ref{scoring}):
%
\begin{equation} \label{expscore}
S(\mu, x) = \la \theta(\mu), \phi(x) \ra - C(\theta(\mu)).
\end{equation}
%
\end{proposition}
%
\noindent
The scoring rule~(\ref{expscore}) decomposes neatly into payoff and cost functions. The first term $\la \theta(\mu), \phi(x) \ra$ defines the agent's outcome-contingent reward, while the second term~$C(\theta(\mu))$ is the cost of acquiring a portfolio $\theta(\mu)$. We see here the duality between the approach of directly reporting $\mu$, or acquiring shares $\theta$: an agent reporting beliefs $\mu$ under scoring rule~(\ref{scoring}) would choose to acquire shares $\theta(\mu)$ in the information market with cost function~(\ref{cost}). Since~(\ref{scoring}) is a proper scoring rule, the information market with cost function~(\ref{cost}) is based on a proper market scoring rule.

\section{The Exponential Family Market Mechanism: A Maximum Likelihood Approach}
In this section, we will set up a prediction market that aggregates beliefs from traders where the outcome is a continuous random variable. In particular, we assume that the outcome is drawn from an exponential family distribution. Each of the traders has access to a series of points drawn from this distribution. In other words, the traders have access an empirical mean of the sufficient statistics of the exponential family. Every trader has infinite budget so that the current market price after a trader has traded in the prediction market, exactly reflects his beliefs.

%We will now provide a more formal description of the problem statement.
%\subsection{Problem Statement}
%\begin{description}
%\item[Setup: ] Learning proceeds in rounds $j=1,2,\ldots$ where at the end of each round the algorithm may be provided with $x_{j}$  drawn from a set $\mathcal{X}$ according to some distribution $\mathcal{D}_{j}$, where $\mathcal{D}_{j}$ belongs to an exponential family defined by $k$ sufficient statistics.  In each round, some sequence of $n$ traders provide their estimates of the empirical means of each of the $k$ sufficient statistics drawn from $\mathcal{D}_{j}$.
%\item[Define: ] A learning algorithm that learns this distribution by providing a maximum likelihood estimate of its unknown natural parameters based on the empirical means.  
%\item[Proposed solution: ] Simulate a prediction market defined by a cost function $C$, securities $s_{0},s_{1},s_{2},\ldots, s_k$ and their payoffs $\phi_i(x)$ for $i=0,1,2,\ldots,k$ for the LMSR. Define a correspondence between the state of the prediction market and the output of the learning algorithm.
%\end{description}

For $x_{i}\sim P_{\betavec}$, recall that the likelihood function for independently drawn data $x_1,\ldots,x_n$ is given by $\prod_{i=1}^{n} P_{\betavec}(x_i)$. The maximum likelihood estimate of the natural parameters $\betavec$ is the value of the natural parameters that maximizes the likelihood function.

We will now set up a prediction market with log market scoring rule (LMSR) and infinite budget traders such that the market state represents the maximum likelihood estimate (MLE) of the natural parameters of an exponential family distribution. %Given  under some interpretation of what information the participating agents have, and how they bet.

For a given exponential family distribution, the prediction market is defined as follows:
\begin{description}
\item[Traders] The prediction market will simulate a trader $i$ corresponding to expert $i$. This trader processes all information samples available to her directly or inferred from previous trades, and  forms a belief distribution such that the believed means of the sufficient statistics matches the empirical means of the sufficient statistics from the information samples. The trader trades in the market to maximize her expected payoff under her believed distribution.
\item[Securities and their payoff]
For each $i=1,2,\ldots,k$, we define a security $s_{i}$ with payoff $\phi_{i}(x)$ where $x$ is the ultimate outcome and $\Phi()$ defines the vector of sufficient statistics of the exponential family distribution according to which $x$ is drawn. We define an additional security $s_{0}$ with payoff $\phi_{0}(x):=a-\sum_{i=1}^{k}\phi_{i}(x)$ where $a$ is an appropriately chosen constant dependent on the range of $\Phi$ so that the payoff of $s_{0}$ is non-negative. 
\end{description}

We also note that since we want non-negative payoffs, we restrict the exponential families under consideration so that the sufficient statistics are lower bounded. If we have a constant lower bound on the sufficient statistics, then without loss of generality we can add a constant to each sufficient statistic without changing the exponential family in any way, as the constants will be absorbed in the log-partition function.

\paragraph{Arbitrage-free Property}
We can easily show that the prediction market we have defined is arbitrage-free: there is no sequence of trades that guarantees the trader a profit under all conditions. First, note that the incremental cost and payoff function are additive over multiple trades, so the net profit of a sequence of trades depends only on the initial and final market position, and is independent of the actual path along which the trade takes place.

>>>>

\section{Maximum Likelihood Scoring Rules}

@@

\begin{itemize}
\item Explain that the scoring rule from Example 2 does not elicit the mean when the agent's distribution is exponential (with change of variable so that it lies in $[1,+\infty)$. 
\item Observe that scoring rule from Example does elicit the mean when the agent's distribution is exponential, Pareto, or in fact any other distribution with support on $[0,+\infty)$ such as the lognormal. 
\item This occurs because the scoring rule is linear in $x$, so its expectation is the same for any distribution with mean $\mu$. 
\item Another feature to note is that under an appropriate parametrization (rate instead of mean), the scoring rule (1) is convex in the parameter being elicited. 
\item These two observations taken together lay the groundwork for the characterization of strongly proper scoring rules for distribution statistics. 
\item Before we proceed should note that the arguments behind Theorem~\ref{maxlike} apply not just to maximum likelihood but to any \emph{m-estimator}, which are notions from robust statistics. Explain what an m-estimator looks like and how there are general consistency results for those as well.
\item The point of these alternative estimators is to be more robust to outliers in the sample, and so these kinds of estimators could translate into scoring rule more robust to inaccuracies in the agent's assessment of the underlying probability distribution or its parameters. (Just a hunch.)
\end{itemize}
